---
title: "Lab 1"
author: "Eduardo Armenta"
date: "2022-08-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### A
It took me 11 total rolls in the simulation to get 3 ones.
6 rolls for the first, 2 for the second, and 3 for the third.

### B
I reused the functions shown to us in the lab diary for die roll and attempt count.
```{r 1}
myroll = function(p){
  u <- runif(1)
  x <- as.numeric(u < p)
  return(x)
}
```
```{r 2}
myattempts = function(p){ counter <- 0
while (myroll(p) == 0){ counter <- counter + 1 }
return(counter) }
```

### C
This function sums the amount of rolls it took to get a one three times.
```{r 3}
sum(replicate(3,myattempts(1/6)))
```

### D
This code simulates the attempt 10,000 times, separates the attempts that took longer than my first one into a vector, counts the vector length and divides it over 10,000 to get the percentage of tries that took longer than the original attempt.
```{r 4}
x <- replicate(10000,sum(replicate(3,myattempts(1/6))))
y <- c()
for (i in x) {
  if (i > 11) {
    y <- c(y, i)
  }
}

percentage = length(y)/100

paste0('The percentage of attempts that did worse than my 11 rolls is ', percentage, '%')
```

## Problem 2

### A: What does it mean to set a seed? How could it be useful to the study of probability and statistics?
Setting a seed allows us to replicate a random event. You select any number you want for the seed to ensure that random events/variables you recreate yield the same results. This is useful for the study of probability and statistics because other it allows other data scientist to recreate your experiment. If there were no seed, generating the same values from a random events would be close to impossible and that would make the experiment difficult to duplicate.

### B: How might you run a similar simulation in python? are there any functions comparable to the ones used in problem #1 and #2?
There's a function in Python from the random module that allows you to set a seed (`random.seed(x)`). You can run the simulation with the same functions as in R, the only difference is that the format for creating them will be slightly different. `int()` ub R is like `as.numeric()` in Python; `random.uniform()` is like `runif()`, etc. I don't know if Python has a function like replicate but you can create a for/while loop that'll run a function 10000 times.

### C: Reflect on your experiences during the boot camp this summer. What might be some advantages/disadvantages of R over Python?
I have a personal preference for Python over R. The syntax feels more natural, but that is likely due to my prior knowledge of the language. I'm assuming each language has certain functions that the other doesn't, which would make me prefer one over the other based on the work I'm doing. That being said, the Rmd files are extremely convenient; I've yet to find  Python's equivalent.
Furthermore, Python has the ability to do more than R. To me, it seems that R is very specific to data science work while Python can be used to create websites, programs, to do data science, and many more things.

## Problem 3
Problem 1.5 in ch. 1 of Dalgaard. On p. 27, replicate was used to simulate the distribution of the mean of 20 random numbers from the exponential distribution by repeating the operation 10 times. That code is:

```{r 5}
set.seed(123)
replicate(10,mean(rexp(20)))
```

How would you do the same thing with `sapply`?

Within the `sapply`, you define how many times you have to apply the function, then you create a function that finds the mean of 20 random exponential distribution numbers.
```{r 6}
set.seed(123)
sapply(1:10, function(i) mean(rexp(20)))
```